---
title: "MC"
author: "Vaclav Haisman"
date: "10 března 2016"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(sqldf)
library(tcltk)
library(gsubfn)
library(proto)
library(RSQLite)
library(DBI)
library(ggplot2)
library(ppls)
```

```{r, echo=FALSE}
years <- rep(2010:2015, each=12)

months <- rep(1:12, 6)

applicants <- c(14970, 16695, 18150, 14490, 14175, 15060, 15475, 17535, 20545, 22270, 20055, 18960,
  17820, 19360, 21955, 19845, 25045, 22935, 22385, 22545, 23975, 23375, 22685, 21300,
  20345, 18915, 19610, 17765, 20175, 20385, 23405, 24080, 26650, 33745, 29515, 23685,
  25840, 23290, 25485, 28825, 29875, 31575, 33625, 31580, 36870, 40745, 35405, 29740,
  35775, 31020, 32925, 34820, 38885, 41420, 51480, 50180, 61755, 62945, 58690, 62780,
  61450, 66150, 61940, 59085, 68090, 90140, 116200, 141250, 165305, 167060, 154605, 104360)

ftaDF <- data.frame(years, months, applicants)

ftaNormDF <- sqldf(
  "SELECT t2.years AS years, t2.months AS months, mi.mi, ma.ma, ((t2.applicants - mi.mi) / (ma.ma - mi.mi)) AS normYearApplicants
    FROM ftaDF t2
    JOIN (SELECT MIN(applicants) AS mi, years FROM ftaDF t2 GROUP BY t2.years) mi USING (years)
    JOIN (SELECT MAX(applicants) AS ma, years FROM ftaDF t2 GROUP BY t2.years) ma USING (years)")
```

# Migration crisis data analysis

## EU-28 first time asylum applicants

This chart shows first time asylum applicants each year since 2010. The data is taken from [Eurostat](http://appsso.eurostat.ec.europa.eu/nui/show.do?query=BOOKMARK_DS-055296_QID_4470A6E0_UID_-3F171EB0&layout=TIME,C,X,0;GEO,L,Y,0;CITIZEN,L,Z,0;SEX,L,Z,1;AGE,L,Z,2;ASYL_APP,L,Z,3;UNIT,L,Z,4;INDICATORS,C,Z,5;&zSelection=DS-055296ASYL_APP,ASY_APP;DS-055296UNIT,PER;DS-055296CITIZEN,EXT_EU28;DS-055296INDICATORS,OBS_FLAG;DS-055296AGE,TOTAL;DS-055296SEX,T;&rankName1=UNIT_1_2_-1_2&rankName2=AGE_1_2_-1_2&rankName3=CITIZEN_1_2_-1_2&rankName4=INDICATORS_1_2_-1_2&rankName5=ASYL-APP_1_2_-1_2&rankName6=SEX_1_2_-1_2&rankName7=TIME_1_0_0_0&rankName8=GEO_1_2_0_1&sortC=ASC_-1_FIRST&rStp=&cStp=&rDCh=&cDCh=&rDM=true&cDM=true&footnes=false&empty=false&wai=false&time_mode=NONE&time_most_recent=false&lang=EN&cfo=%23%23%23%2C%23%23%23.%23%23%23).

```{r, echo=FALSE, fig.align='center', fig.height=3}
ggplot(ftaDF, aes(x = factor(months), y = applicants / 1000, group = years, color = factor(years))) +
  geom_point() +
  geom_line() +
  labs(color = "years", x = "month", y = "thousands of first time applicants",
       title = "EU-28 first time asylum applicants")
```

In the following chart, the values are normalized per year so that we can see annual cycles in migration. In years 2012-2015, the migration swells in the second half of the year and falls off significantly in December.

```{r, echo=FALSE, fig.align='center', fig.height=3}
ggplot(ftaNormDF, aes(x = factor(months), y = normYearApplicants, group = years, color = factor(years))) +
  geom_point() +
  geom_line() +
  labs(color = "years", x = "month", y = "per year normalized applicants",
       title = "EU-28 first time asylum applicants (per year normalized)")
```
